---
title: "Video Games, Self-Concept, and the Parent–Child Relationship"
subtitle: "Analysis report"
author: "Taren Sanders"
date: today
# The docx is the reviewed artifact; the html is a quick-preview convenience
# (first so the VS Code Quarto preview uses it). Both render in the pipeline.
format:
  html:
    toc: true
    number-sections: true
    embed-resources: true
  docx:
    toc: true
    number-sections: true
    reference-doc: reference-portrait.docx
execute:
  echo: false
  warning: false
  message: false
---

```{r setup}
library(targets)
library(ggplot2)
library(knitr)

# Quarto renders from the document's directory, so point tar_read() at the
# project-root store explicitly.
store <- here::here("_targets")
```

# Methods

- Data: LSAC B cohort at ages 10/11, 12/13, and 14/15 (waves 6–8).
  The K cohort is excluded because its age-14 wave (2012) did not have the video game time item.

## Key Variables

- Video game use: reported minutes of electronic games in an average week, converted to hours/week.
- Self-concept: self-reported SDQ.
  The total difficulties score is the primary outcome, and internalising (emotional + peer) and externalising (conduct + hyperactivity) composites are for secondary analyses.
  The Parent 1-reported SDQ total is also kept for an informant-robustness check.
- Parental warmth (the moderator): Parent 1's score on the 6-item Parental Warmth Scale (1–5).
  Parent 2's score is retained for a both-parents sensitivity model.
  It's not the primary model because so much of Parent 2's data is missing.
- Covariates: sex, socioeconomic position, and age in months, all at the age 10/11 baseline.
- Cleaning: LSAC missing-data codes recoded to missing.
  Video game hours more than 4 SD above the within-wave mean are winsorised (that is, set to the 4 SD value).
  That is different from the proposal (which said set to NA), but I think makes more sense.
  The flagged values look like real heavy gamers overstating, so we keep them but cap their influence.
  I'll also test setting them to missing as a sensitivity analysis.
  SES and warmth are not screened — there are only a few extreme SES values and these seem plausible given the size of LSAC, and low warmth scores are likely to be real given that these are likert scale items.
- Analysis sample: all children with data on at least one study variable (video games, SDQ, or warmth) at one or more waves, with models estimated using full-information maximum likelihood (FIML) under a missing-at-random (MAR) assumption.
  The proposal's complete-case (data at all three waves) can be done as a sensitivity analysis.

# Data and derivations

## Sample and missingness per wave

```{r tbl-missingness}
knitr::kable(
  tar_read(missingness_table, store = store),
  caption = "Non-missing observations (% of children in the wave file) for each study variable, after cleaning. N is the number of children with any data at that wave."
)
```

## Distributions

```{r plot-distributions, fig.cap="Distributions of the three core study variables at each age band (after cleaning).", fig.width=8, fig.height=7}
tar_read(distributions_plot, store = store)
```

## Video game hours: outlier tail {#sec-vg-tail}

Values more than 4 SD above the within-wave mean are winsorised to 4 SD from mean.
The flagged values are shown below before windsorising.

```{r tbl-vg-tail}
knitr::kable(
  tar_read(vg_tail_table, store = store),
  caption = "Unscreened video game hours per week: within-wave summary and all values above the 4 SD flagging threshold."
)
```

# Sample and descriptives

## Sample flow

```{r sample-flow}
sample_flow <- tar_read(sample_flow, store = store)
n_file <- sample_flow$flow$`N children`[1]
n_analysis <- sample_flow$flow$`N children`[2]
n_complete <- sample_flow$flow$`N children`[7]
```

The three B-cohort wave files contain `r format(n_file, big.mark = ",")` children.
The analysis sample (children with data on at least one study variable at one or more waves) is N = `r format(n_analysis, big.mark = ",")`.
The proposal's complete-case criterion would retain N = `r format(n_complete, big.mark = ",")` (`r sample_flow$flow[["% of cohort file"]][7]`% of the cohort file).
That makes me think that FIML is definitely the way to go.

```{r tbl-sample-flow}
knitr::kable(
  sample_flow$flow,
  caption = "Sample flow from the wave 6–8 B-cohort files to the analysis samples."
)
```

```{r tbl-participation-patterns}
knitr::kable(
  sample_flow$patterns,
  caption = "Wave-participation patterns (any study data at each wave) among the analysis sample.",
  align = c("c", "c", "c", "r", "r")
)
```

::: {.callout-note title="Interpretation"}
- Three-quarters of the analysis sample (74.2%) have study data at all three waves, and 86.6% at two or more.
- Complete-case analysis would discard more than a third (36.9%) of the sample.
  So FIML is a good option.
:::

## Descriptive statistics

```{r descriptives}
descriptives <- tar_read(descriptives_table, store = store)
knitr::kable(
  descriptives$by_age,
  caption = "Descriptive statistics by age band. N per column is children with any data at that wave; individual variables have additional item-level missingness."
)
```

## Trajectories

```{r plot-trajectories, fig.cap="Video game hours and self-reported SDQ total difficulties by age band: distributions (violins) with means and 95% CIs.", fig.width=8, fig.height=4}
tar_read(trajectory_plot, store = store)
```

::: {.callout-note title="Interpretation"}
- Video game use rises with age (7.7 → 8.5 → 10.1 hrs/week on average) and, quite interestingly, spreads out.
  The SD grows from 6.7 to 11.2, with a lengthening upper tail.
- SDQ total difficulties dips at 12/13 (10.1 → 9.3) then returns to baseline at 14/15 (10.3).
  It's pretty small, so I wouldn't overinterpret it.
- Parental warmth declines steadily with age (4.35 → 4.08).
  Again pretty small, and might just be due to dropout?
:::

# Missingness and attrition {#sec-attrition}

Children with any study data at age 10/11 who also have study data at either follow-up wave ("retained") are compared with those who have none ("lost") on their baseline characteristics.
Standardised mean differences (SMD) are the retained-minus-lost difference in pooled-SD units.
|SMD| < 0.10 is conventionally negligible.

```{r tbl-attrition}
knitr::kable(
  tar_read(attrition_table, store = store),
  caption = "Baseline (age 10/11) characteristics of children retained at follow-up vs lost to follow-up."
)
```

::: {.callout-note title="Interpretation"}
- Attrition is clearly not completely at random: lost children are lower-SES (SMD 0.37) and report slightly higher baseline difficulties (SMD −0.12).
  Sex, warmth, and video game hours show negligible differences.
- The variables that predict dropout (SES, baseline SDQ) are observed and in the model.
  This is the MAR pattern FIML adjusts for, whereas complete-case analysis would tilt the sample toward higher-SES, better-functioning children.
:::

# Correlations

```{r tbl-correlations}
knitr::kable(
  tar_read(correlation_table, store = store),
  caption = "Zero-order correlations among the study variables (pairwise complete). SES is the age-10/11 z-composite; Female sex is coded 1 = female. \\* p < .05, \\*\\* p < .01, \\*\\*\\* p < .001."
)
```

::: {.callout-note title="Interpretation"}
- The cross-sectional and cross-lagged video-game–SDQ correlations are all small and positive (r = .04–.17).
  - More gaming goes with slightly more difficulties, before any adjustment.
- Rank-order stability is moderate for video games (.42–.51), higher for SDQ (.57–.61) and warmth (.69–.70).
- Warmth correlates negatively with both gaming and difficulties (r ≈ −.05 to −.12).
- Sex is strongly related to gaming (girls lower, and increasingly so with age: −.24 → −.48) and flips sign for SDQ (−.09 at 10/11 to +.14 at 14/15).
   — Supports sex as a covariate in all models.
:::

# Cross-lagged models (H1)

```{r model-fit-setup}
model_fit <- tar_read(model_fit_table, store = store)
n_sem <- model_fit$N[model_fit$Model == "RI-CLPM (primary)"]
```

## What the models do

- Both models estimate whether video game use predicts later SDQ total difficulties (H1a) and whether difficulties predict later video game use (H1b), across the three age bands.
- The traditional CLPM relates each child's score to their earlier score without separating "this child games more than other children" from "this child is gaming more than they usually do".
  Sex and socioeconomic position are included as covariates.
- The RI-CLPM adds a random intercept for each variable, which captures each child's stable average level.
  The lagged paths are then estimated on what is left over (deviations from a child's own average) so they are the within-person test.
  I've used this one as the primary model.
- The RI-CLPM does not include sex or SES.
  The random intercepts already absorb every stable difference between children, including those two and anything unmeasured, so the lagged paths are not confounded by them.
- Both are estimated with robust maximum likelihood (RML) and FIML, on the `r format(n_sem, big.mark = ",")` children with data on at least one of the six modelled variables.
- I ran everything for both types of models, so the choice of primary model can be changed without redoing the analysis.

## Within-child and between-child {#sec-within-between}

The two models answer different questions.

- A **between-child** question compares children to each other: *do children who game more than other children report more difficulties than other children?*
- A **within-child** question compares a child to themselves at other times: *when a child games more than is usual for them, do they report more difficulties than is usual for them, later on?*

I think the RI-CLPM makes sense for the primary model because the hypotheses are about the future (e.g., H1a is that gaming predicts *subsequent* self-concept), and a between-child comparison does not really address that on its own.

A child in a chaotic household might game more and report more difficulties at every wave, which produces an association with no causal link between the two.
The random intercepts help with this.

## Model fit

```{r tbl-model-fit}
knitr::kable(
  model_fit,
  caption = "Fit of the cross-lagged models. Robust (MLR-corrected) indices. Fit statistics and information criteria are only comparable between models fitted to the same variables, so the covariate-adjusted CLPM (8 variables) cannot be compared with the other two (6 variables)."
)
```

::: {.callout-note title="Interpretation"}
- The RI-CLPM fits substantially better than the CLPM on the same six variables: RMSEA .072 vs .115, CFI .990 vs .947, and a BIC around 157 points lower.
- Most of what looks like "gaming predicts difficulties" in the CLPM is stable between-child difference.
  Some children game more and report more difficulties than others, rather than change within a child over time.
- No model reaches a non-significant χ², but with N ≈ 3,900 that is expected and not very informative.
:::

## Does the effect stay the same across the two lags? {#sec-h1-decisions}

The model can either estimate each lag separately or assume the paths are the same size from 10/11 → 12/13 as from 12/13 → 14/15.
Assuming they are equal is normally preferred with only three waves, because leaving everything free leaves just one degree of freedom.

```{r tbl-stationarity}
stationarity <- tar_read(stationarity_riclpm, store = store)
knitr::kable(
  stationarity$comparisons,
  caption = "Scaled chi-square difference tests of the constrained RI-CLPMs against the freely estimated model."
)
```

```{r tbl-stationarity-paths}
knitr::kable(
  stationarity$constraints,
  caption = "Univariate score tests on the fully constrained RI-CLPM: which lagged paths actually differ between the two lags."
)
```

The score tests above are computed with the autoregressive equality constraints still imposed, and those are violated.
The following asks the same question of the freely estimated model, showing each lag's own estimate next to the test of whether they differ.

```{r tbl-cross-lag-equality}
knitr::kable(
  stationarity$cross_lags,
  caption = "Cross-lagged paths estimated separately at each lag, with a Wald test of whether they differ, from the freely estimated RI-CLPM."
)
```

::: {.callout-note title="Interpretation"}
- Assuming all paths are equal across lags is clearly wrong here.
- But the problem is mostly in the autoregressive paths (how strongly a child's own earlier deviation predicts their later one) not in the cross-lagged paths.
  Since the hypotheses are about the cross-lagged paths it makes sense to just focus on those.
- The two hypotheses differ in how safely they can be pooled.
  H1b is always stable: .111 at the first lag and .093 at the second and statistically not different (p = .538).
  H1a is not stable: .101 at the first lag and −.007 at the second, and borderline significant (p = .049).
- The joint test of both cross-lags together is not rejected (p = .099), but that is probably because H1b is stable and that dilutes H1a's instability.
- So the primary model pools the cross-lags (giving one estimate per hypothesis) and lets the autoregressions differ.
  That version is not significantly worse than the free model, and fits better than either alternative but the pooled H1a estimate should be read with the two lags above in mind.
- I checked whether this was caused by video game hours being skewed, by refitting on square-root and log-transformed hours, and the estimates are all basically the same.
  My guess is that the cause is the spread of gaming nearly triples between 10/11 and 14/15 (variance 45 → 125) as more children stop playing entirely and others play much more, and transformaing the variables doesn't fix that (on purpose - we want to keep variance).
:::

## Within-child effects: RI-CLPM (primary)

```{r tbl-riclpm}
knitr::kable(
  tar_read(riclpm_table, store = store),
  caption = "RI-CLPM structural estimates. Unstandardised coefficients are in hours/week and SDQ points; β are fully standardised. Cross-lagged paths are constrained equal across lags."
)
```

```{r fig-path-diagram, fig.cap="RI-CLPM with standardised estimates. Squares are observed variables, ellipses are latent: the random intercepts capture stable between-child differences, and the within-person components (VG, SDQ) capture each child's deviation from their own average. Blue arrows are the estimated structural paths; grey dashed arrows are fixed at 1. * p < .05, ** p < .01, *** p < .001.", fig.width=8, fig.height=6.5}
tar_read(path_diagram, store = store)
```

::: {.callout-note title="Interpretation"}
- OK, the actual models.
  Here's how I would read these:
  - H1b is supported. 
    When a child reports more difficulties than usual for them, they report more video game use than usual two years later (β = .091, 95% CI [.024, .158], p = .006).
    The effect is small: roughly 0.13 extra hours per week per additional SDQ point.
  - H1a is not supported.
    Pooled across both lags, gaming does not predict later difficulties (β = .004, 95% CI [−.026, .034], p = .783).
    That pooled estimate averages a marginal positive effect from 10/11 to 12/13 (β = .101, p = .054) and a null from 12/13 to 14/15 (β = −.007, p = .785), and those two just barelydiffer from each other (p = .049).
    The narrow pooled confidence interval is partly a consequence of pooling, and should not be read as ruling out an effect at every age.
- Another summary of H1a is "null on average, with a marginal early effect that has gone by 14/15" rather than "always null".
  It doesn't really matter though.
  The first-lag effect is only p = .054 even when estimated freely, so the effect would not have been significant if estimated freely.
- So the relationship looks one-directional, and in the opposite direction to what we probably would have guessed: difficulties appear to lead gaming rather than gaming leading difficulties.
- The random intercepts correlate .21 (p < .001).
  Children who game more on average do report more difficulties on average, but that is a between-child difference, not evidence that gaming changes anything.
- One thing I did find hard to explain: the within-person video game autoregression is −.26 at the first lag and +.28 at the second.
  A negative within-person autoregression is not impossible (a spike above your own average tends to be followed by a dip below it), but a sign flip that large is more likely instability from the three-wave model combined with gaming's growing variance.
  It does not affect the cross-lagged conclusions, which are stable across every specification tested, but we should keep it in mind.
:::

## Between-child differences {#sec-between-person}

The same RI-CLPM also estimates the between-child side of the relationship, in the correlation between the two random intercepts.
That is, between a child's stable average gaming level and their stable average level of difficulties.
This is the answer to "do children who game more than other children report more difficulties than other children?".

```{r between-person}
riclpm_estimates <- tar_read(riclpm_table, store = store)
ri_row <- riclpm_estimates[
  riclpm_estimates$Parameter == "Random intercept correlation",
]
```

Children who game more on average do report more difficulties on average: r = `r ri_row[["β"]]`, 95% CI `r ri_row[["β 95% CI"]]`, p `r ri_row[["p"]]`.

This between-child association is larger than either within-child cross-lagged effect in the model.
Most of the raw relationship between gaming and difficulties (which is in the correlation table above) is of this kind: a difference between children rather than a process operating within them.

It's usefil to look at how much of each variable exists at each level:

```{r tbl-variance-split}
knitr::kable(
  tar_read(variance_split, store = store),
  caption = "Proportion of each variable's variance that is stable between-child difference versus within-child variation, from the RI-CLPM."
)
```

::: {.callout-note title="Interpretation"}
- Difficulties are consistently about 40% between-child at every age: some children simply report more problems than others, stably, and that share does not change as they get older.
- Video game use behaves very differently.
  At 10/11 it is strongly trait-like (68% between-child), but by 14/15 only 25% is stable difference and three quarters is time-specific.
  Gaming stops being "the sort of child you are" and becomes "what you were doing that year".
- That shift is the same phenomenon as the growing variance noted earlier, and it is the most likely explanation for the unstable within-child gaming autoregression: there is progressively less stable signal to carry forward.
- Putting it all together: there is a real and reasonably clear association between gaming and difficulties **between** children (r = .21), and essentially no evidence that gaming drives difficulties **within** a child over time (β = .004).
:::

## Comparison model: CLPM estimates

This is mostly a repeat of the analysis above, but using a 'classic' CLPM instead of the RI-CLPM.

```{r tbl-clpm}
knitr::kable(
  tar_read(clpm_table, store = store),
  caption = "Traditional CLPM structural estimates, adjusted for sex and socioeconomic position. Cross-lagged paths are constrained equal across lags."
)
```

```{r fig-clpm-path-diagram, fig.cap="Traditional CLPM with standardised estimates. All variables are observed (squares); with no random intercepts, the lagged paths run directly between them and mix between-child and within-child variation. Sex and socioeconomic position are included as covariates but their paths are omitted for legibility. * p < .05, ** p < .01, *** p < .001.", fig.width=8, fig.height=3.6}
tar_read(clpm_path_diagram, store = store)
```

::: {.callout-note title="Interpretation"}
- The CLPM finds both cross-lagged paths significant (H1a β = .051, H1b β = .082, both p < .001), which would support a bidirectional relationship.
- The RI-CLPM shows that the H1a path is an artefact of not separating between- from within-child variance.
- I'd suggest you go and read [Hamaker et al., 2015](https://doi.org/10.1037/a0038889) which is where the RI-CLPM comes from and it explains exactly this kind of problem.
  It is a bit of a dense read though.
- Reporting the CLPM alone would have produced a different (and I think wrong) conclusion.
:::

# Moderation by parenting (H2)

## What the test does

- H2 is that a better parent–child relationship weakens the cross-lagged associations, that gaming would matter less for children who feel closer to their parents.
- In the protocol we just said we would use parental warmth, but since I had the other variables I tested them too.
- I don't think you should report all of this.
  Since there is nothing interesting, I would stick with just the variable you prespecified (warmth).
  You could maybe mention that you tested other parenting constructs as sensitivity analyses, but it just means more for you to write up.
- It's worth knowing that the five parenting constructs tested come from two different informants.
  Warmth and angry parenting are the parent's own report.
  Responsiveness, autonomy-granting and demandingness come  from the child's own questionnaire, rating their mother and their father separately. 
  There is no parent-report version.
  The mother version is used, as it has the best coverage and adding the father version gains only three children.
- All five are scored so that higher means more of the named quality.
- Each test splits children into the lowest and highest thirds of that construct at age 10/11 and asks whether the two cross-lagged paths are the same size in both groups.
  If the construct moderates, forcing the paths equal across groups should worsen fit.
- Grouping is at baseline only, so each child stays in one group across all three waves.
  That assumes parenting is pretty stable, which I think is pretty reasonable assumption.
- Everything else in the model (autoregressions, variances, random intercepts) is free to differ between groups, so the test isolates the paths H2 is about.
- Both model families are run, as for H1.

```{r moderation-setup}
moderation <- tar_read(moderation_test, store = store)
```

## Does any parenting construct change the cross-lagged paths?

```{r tbl-moderation-omnibus}
knitr::kable(
  moderation$omnibus,
  caption = "Test of H2 for each parenting construct: scaled chi-square difference between models with the cross-lagged paths free to differ between the lower and upper tertile groups and constrained equal across them."
)
```

::: {.callout-note title="Interpretation"}
- H2 is not supported for any of the five constructs, in either model family.
  All ten omnibus tests are non-significant, and most are far smaller than their two degrees of freedom (which is pretty clear evidence of nothing happening).
- The one exception is autonomy-granting in the CLPM, at p = .050 (exactly .0504, so technically above 0.05).
  Its RI-CLPM counterpart is null (p = .149).
  With ten tests run, one significant result is basically what you expect based on chance alone so I wouldn't read anything into it.
- Also the autonomy RI-CLPM's constrained model returns a negative variance, so that particular comparison is almost certainly wrong.
:::

```{r tbl-moderation-paths}
knitr::kable(
  moderation$paths,
  caption = "Wald tests of each cross-lagged path separately across the tertile groups of each construct."
)
```

## Estimates by group

```{r fig-moderation, fig.cap="Standardised cross-lagged estimates with 95% confidence intervals, by baseline tertile group on each parenting construct and by model family. H2 predicts the amber intervals sitting to the left of the blue ones.", fig.width=8, fig.height=7}
tar_read(moderation_plot, store = store)
```

```{r tbl-moderation-estimates}
knitr::kable(
  tar_read(moderation_table, store = store),
  caption = "Cross-lagged estimates by tertile group for each parenting construct. Cross-lagged paths are constrained equal across the two lags but free to differ between groups."
)
```

::: {.callout-note title="Interpretation"}
- The confidence intervals for the two groups overlap in essentially every panel, so visually they look null too.
- Splitting the sample roughly halves the power for each group, so the H1b effect that was significant in the full sample (β = .091, p = .006) is often not significant within a single tertile.
  That is reduced precision from subgrouping, not a contradiction of the H1 result.
:::

## Sensitivity: the moderators as continuous variables {#sec-h2-continuous}

The multi-group tests categorise each construct and drop the middle third.
These models keep each one continuous, use the whole sample, and test moderation as an interaction term instead.
They are fitted as CLPMs only: the RI-CLPM equivalent would need a latent interaction (which is tricky and probably non-sensical in three waves of data).

```{r tbl-interaction}
knitr::kable(
  tar_read(interaction_table, store = store),
  caption = "Continuous moderation models: interaction between each parenting construct at baseline and the predictor at each lag. A significant interaction would indicate moderation."
)
```

::: {.callout-note title="Interpretation"}
- The continuous tests agree with the multi-group tests: no construct moderates either direction.
:::

# Sensitivity and secondary analyses

## What these check

Each analysis below changes exactly one decision and leaves everything else as specified, so any change in the results can be attributed to that decision.
All are run in both model families.

- **Complete cases only** — the proposal's inclusion criterion, against the FIML approach adopted instead.
- **SDQ internalising / externalising** — the two broad composites in place of total difficulties.
- **Parent-reported SDQ** — the same models with the parent's rating instead of the child's.
- **Outliers set to missing** — the alternative to winsorising heavy gaming.
- **Log video game hours** — the transformed scale.

```{r tbl-sensitivity}
knitr::kable(
  tar_read(sensitivity_table, store = store),
  caption = "Cross-lagged paths across the sensitivity analyses. β with 95% CI; \\* p < .05, \\*\\* p < .01, \\*\\*\\* p < .001. \"Outcome\" is SDQ total difficulties unless the row says otherwise."
)
```

::: {.callout-note title="Interpretation"}
- The H1a null is very robust. In the RI-CLPM, gaming fails to predict later difficulties under every variation using child-reported SDQ (complete cases, both subscales, either outlier rule, either scale).
- The H1b is fairly robust. It holds for the primary specification (β = .091), externalising (β = .094) and the alternative outlier rule (β = .085), but falls short of significance under complete cases (β = .066), internalising (β = .048) and log hours (β = .038).
  Both of these are explainable though.
  The complete-case drops >1,400 children (so lose power), and log-transforming compresses the kids who game the most (and presumably carry most of the signal).
- Internalising and externalising behave differently. The H1b effect is carried by externalising difficulties (conduct and hyperactivity), not internalising (emotional and peer).
  We didn't prespecify this, so I'm reluctant to dig into it too much, but we can if you want.
- The CLPM is more robust to everything. It returns both paths significant in all seven variations.
  A model that gives the same answer regardless of specification is not necessarily more trustworthy than one that responds to the data.
:::

## Informant matters {#sec-informant}

Switching from the child's own SDQ to the parent's rating changes the results more than every other sensitivity analysis.
In the RI-CLPM, H1a goes from null (β = .004) to significant (β = .060, p < .001), and H1b roughly doubles (β = .091 to β = .216).

The two informants agree only moderately (child and parent SDQ correlate .43 to .46 across the three waves) so they are not interchangeable measures.
Probably more tellingly, video game hours correlate more strongly with the parent's rating of difficulties than with the child's, and it gets worse with age:

| Age band | VG with child-reported SDQ | VG with parent-reported SDQ |
| --- | --- | --- |
| 10/11 | .120 | .141 |
| 12/13 | .132 | .200 |
| 14/15 | .036 | .188 |

::: {.callout-note title="Interpretation"}
- By 14/15 gaming is essentially unrelated to how children rate their own difficulties (r = .04) but clearly related to how their parents rate them (r = .19).
- Shared-informant variance is the leading candidate explanation: if the same parent reports both the child's gaming and their difficulties, correlated reporting bias inflates the association between them.
- It could also be substantive: parents may notice difficulties in heavy-gaming adolescents that the adolescents do not report in themselves, or parental concern about gaming may influence their ratings of the child generally.
- Either way, the conclusion drawn depends heavily on whose report is used.
  You should probably mention this.
:::

## Between-child differences {#sec-supp-between}

The primary model deliberately omits sex and SES because its random intercepts already absorb them.
This asks the separate descriptive question of which children sit high on the stable levels, using each child's average across the waves they have data for.

```{r tbl-between-person}
knitr::kable(
  tar_read(between_table, store = store),
  caption = "Sex and socioeconomic position as predictors of each child's average video game use and average SDQ total difficulties."
)
```

::: {.callout-note title="Interpretation"}
- Sex dominates gaming: boys average about 6.2 more hours per week than girls (β = −.44), and sex plus SES together explain 21% of the variance in average gaming.
- Sex is unrelated to average difficulties (p = .727), while lower SES predicts both more gaming and more difficulties.
  Since SES predicts both, it is a plausible common cause of the between-child association and the kind of confounding the random intercepts remove from the within-child estimates.
- These are descriptive associations about who children are, not about what gaming does.
:::

The originally planned version of this analysis (regressing the RI-CLPM's random intercepts on sex and SES) does not produce a usable solution:

```{r tbl-between-diagnostics}
knitr::kable(
  tar_read(between_diagnostics, store = store),
  caption = "Diagnostics for the planned random-intercept version of the between-child model."
)
```

Both random-intercept variances are estimated as negative, with standard errors larger than the estimates themselves, and standardised estimates cannot be computed at all.
This is an improper solution so its coefficients are not interpretable.
The person-mean analysis above answers the same question anyway.

## Exploratory: does sex moderate the cross-lagged paths? {#sec-supp-sex}

This is just me mucking around, not something we said we would do and probably not something you need to report.
Given how strongly sex relates to gaming, it is worth checking whether the processes differ for boys and girls.

```{r tbl-sex-moderation}
sex_moderation <- tar_read(sex_moderation_test, store = store)
knitr::kable(
  sex_moderation$omnibus,
  caption = "Exploratory test of whether the cross-lagged paths differ by sex."
)
```

```{r tbl-sex-moderation-paths}
knitr::kable(
  sex_moderation$paths,
  caption = "Wald tests of each cross-lagged path separately across sex."
)
```

```{r fig-sex-moderation, fig.cap="Standardised cross-lagged estimates with 95% confidence intervals, by sex and model family.", fig.width=8, fig.height=3.4}
tar_read(sex_moderation_plot, store = store)
```

::: {.callout-note title="Interpretation"}
- No evidence that either cross-lagged path differs between boys and girls (RI-CLPM Δχ²(2) = 0.06, p = .971; CLPM Δχ²(2) = 0.23, p = .890).
- Sex therefore matters a great deal for how much children game, and not at all for how gaming and difficulties relate to each other over time.
:::

# Supplementary material

## Model fit for the moderation models {#sec-supp-moderation-fit}

```{r tbl-moderation-fit}
knitr::kable(
  tar_read(moderation_fit_table, store = store),
  caption = "Fit of the multi-group moderation models. N is the combined sample across both warmth groups."
)
```

## Stationarity in the CLPM {#sec-supp-stationarity-clpm}

The same constraint tests for the comparison CLPM, for completeness.

```{r tbl-stationarity-clpm}
stationarity_clpm <- tar_read(stationarity_clpm, store = store)
knitr::kable(
  stationarity_clpm$comparisons,
  caption = "Scaled chi-square difference tests of the constrained CLPMs against the freely estimated CLPM."
)
```

```{r tbl-stationarity-clpm-paths}
knitr::kable(
  stationarity_clpm$constraints,
  caption = "Univariate score tests on the fully constrained CLPM."
)
```

The CLPM does not show the same pattern as the RI-CLPM.
Full stationarity is rejected here too, but so is the cross-lags-only constraint.

The CLPM reported in the main text still uses the same constraint as the RI-CLPM, so that the fit comparison in @tbl-model-fit is like-for-like.
It does not change what the CLPM concludes: with all lagged paths freely estimated its cross-lags remain significant in both directions and at both lags (video games → SDQ β = .072 and .043; SDQ → video games β = .055 and .091; all p < .01).
The divergence from the RI-CLPM in @tbl-riclpm is probably driven by the between/within separation, not by the constraint choice.

## Descriptive statistics by sex {#sec-supp-descriptives}

```{r tbl-by-sex-age}
knitr::kable(
  descriptives$by_sex_age,
  caption = "Descriptive statistics by sex within age band."
)
```

