---
title: "Video Games, Self-Concept, and the Parent–Child Relationship"
subtitle: "Analysis report"
author: "Taren Sanders"
date: today
# The docx is the reviewed artifact; the html is a quick-preview convenience
# (first so the VS Code Quarto preview uses it). Both render in the pipeline.
format:
  html:
    toc: true
    number-sections: true
    embed-resources: true
  docx:
    toc: true
    number-sections: true
execute:
  echo: false
  warning: false
  message: false
---

```{r setup}
library(targets)
library(ggplot2)
library(knitr)

# Quarto renders from the document's directory, so point tar_read() at the
# project-root store explicitly.
store <- here::here("_targets")
```

# Methods

- Data: LSAC B cohort at ages 10/11, 12/13, and 14/15 (waves 6–8).
  The K cohort is excluded because its age-14 wave (2012) did not have the video game time item.

## Key Variables

- Video game use: reported minutes of electronic games in an average week, converted to hours/week.
- Self-concept: self-reported SDQ.
  The total difficulties score is the primary outcome, and internalising (emotional + peer) and externalising (conduct + hyperactivity) composites are for secondary analyses.
  The Parent 1-reported SDQ total is also kept for an informant-robustness check.
- Parental warmth (the moderator): Parent 1's score on the 6-item Parental Warmth Scale (1–5).
  Parent 2's score is retained for a both-parents sensitivity model.
  It's not the primary model because so much of Parent 2's data is missing.
- Covariates: sex, socioeconomic position, and age in months, all at the age 10/11 baseline.
- Cleaning: LSAC missing-data codes recoded to missing.
  Video game hours more than 4 SD above the within-wave mean are winsorised (that is, set to the 4 SD value).
  That is different from the proposal (which said set to NA), but I think makes more sense.
  The flagged values look like real heavy gamers overstating, so we keep them but cap their influence.
  I'll also test setting them to missing as a sensitivity analysis.
  SES and warmth are not screened — there are only a few extreme SES values and these seem plausible given the size of LSAC, and low warmth scores are likely to be real given that these are likert scale items.
- Analysis sample: all children with data on at least one study variable (video games, SDQ, or warmth) at one or more waves, with models estimated using full-information maximum likelihood (FIML) under a missing-at-random (MAR) assumption.
  The proposal's complete-case (data at all three waves) can be done as a sensitivity analysis.

# Data and derivations

## Sample and missingness per wave

```{r tbl-missingness}
knitr::kable(
  tar_read(missingness_table, store = store),
  caption = "Non-missing observations (% of children in the wave file) for each study variable, after cleaning. N is the number of children with any data at that wave."
)
```

## Distributions

```{r plot-distributions, fig.cap="Distributions of the three core study variables at each age band (after cleaning).", fig.width=8, fig.height=7}
tar_read(distributions_plot, store = store)
```

## Video game hours: outlier tail {#sec-vg-tail}

Values more than 4 SD above the within-wave mean are winsorised to 4 SD from mean.
The flagged values are shown below before windsorising.

```{r tbl-vg-tail}
knitr::kable(
  tar_read(vg_tail_table, store = store),
  caption = "Unscreened video game hours per week: within-wave summary and all values above the 4 SD flagging threshold."
)
```

# Sample and descriptives

## Sample flow

```{r sample-flow}
sample_flow <- tar_read(sample_flow, store = store)
n_file <- sample_flow$flow$`N children`[1]
n_analysis <- sample_flow$flow$`N children`[2]
n_complete <- sample_flow$flow$`N children`[7]
```

The three B-cohort wave files contain `r format(n_file, big.mark = ",")` children.
The analysis sample (children with data on at least one study variable at one or more waves) is N = `r format(n_analysis, big.mark = ",")`.
The proposal's complete-case criterion would retain N = `r format(n_complete, big.mark = ",")` (`r sample_flow$flow[["% of cohort file"]][7]`% of the cohort file).
That makes me think that FIML is definitely the way to go.

```{r tbl-sample-flow}
knitr::kable(
  sample_flow$flow,
  caption = "Sample flow from the wave 6–8 B-cohort files to the analysis samples."
)
```

```{r tbl-participation-patterns}
knitr::kable(
  sample_flow$patterns,
  caption = "Wave-participation patterns (any study data at each wave) among the analysis sample.",
  align = c("c", "c", "c", "r", "r")
)
```

::: {.callout-note title="Interpretation"}
- Three-quarters of the analysis sample (74.2%) have study data at all three waves, and 86.6% at two or more.
- Complete-case analysis would discard more than a third (36.9%) of the sample.
  So FIML is a good option.
:::

## Descriptive statistics

```{r descriptives}
descriptives <- tar_read(descriptives_table, store = store)
knitr::kable(
  descriptives$by_age,
  caption = "Descriptive statistics by age band. N per column is children with any data at that wave; individual variables have additional item-level missingness."
)
```

## Trajectories

```{r plot-trajectories, fig.cap="Video game hours and self-reported SDQ total difficulties by age band: distributions (violins) with means and 95% CIs.", fig.width=8, fig.height=4}
tar_read(trajectory_plot, store = store)
```

::: {.callout-note title="Interpretation"}
- Video game use rises with age (7.7 → 8.5 → 10.1 hrs/week on average) and, quite interestingly, spreads out.
  The SD grows from 6.7 to 11.2, with a lengthening upper tail.
- SDQ total difficulties dips at 12/13 (10.1 → 9.3) then returns to baseline at 14/15 (10.3).
  It's pretty small, so I wouldn't overinterpret it.
- Parental warmth declines steadily with age (4.35 → 4.08).
  Again pretty small, and might just be due to dropout?
:::

# Missingness and attrition {#sec-attrition}

Children with any study data at age 10/11 who also have study data at either follow-up wave ("retained") are compared with those who have none ("lost") on their baseline characteristics.
Standardised mean differences (SMD) are the retained-minus-lost difference in pooled-SD units.
|SMD| < 0.10 is conventionally negligible.

```{r tbl-attrition}
knitr::kable(
  tar_read(attrition_table, store = store),
  caption = "Baseline (age 10/11) characteristics of children retained at follow-up vs lost to follow-up."
)
```

::: {.callout-note title="Interpretation"}
- Attrition is clearly not completely at random: lost children are lower-SES (SMD 0.37) and report slightly higher baseline difficulties (SMD −0.12).
  Sex, warmth, and video game hours show negligible differences.
- The variables that predict dropout (SES, baseline SDQ) are observed and in the model.
  This is the MAR pattern FIML adjusts for, whereas complete-case analysis would tilt the sample toward higher-SES, better-functioning children.
:::

# Correlations

```{r tbl-correlations}
knitr::kable(
  tar_read(correlation_table, store = store),
  caption = "Zero-order correlations among the study variables (pairwise complete). SES is the age-10/11 z-composite; Female sex is coded 1 = female. \\* p < .05, \\*\\* p < .01, \\*\\*\\* p < .001."
)
```

::: {.callout-note title="Interpretation"}
- The cross-sectional and cross-lagged video-game–SDQ correlations are all small and positive (r = .04–.17).
  - More gaming goes with slightly more difficulties, before any adjustment.
- Rank-order stability is moderate for video games (.42–.51), higher for SDQ (.57–.61) and warmth (.69–.70).
- Warmth correlates negatively with both gaming and difficulties (r ≈ −.05 to −.12).
- Sex is strongly related to gaming (girls lower, and increasingly so with age: −.24 → −.48) and flips sign for SDQ (−.09 at 10/11 to +.14 at 14/15).
   — Supports sex as a covariate in all models.
:::

# Supplementary material

## Descriptive statistics by sex {#sec-supp-descriptives}

```{r tbl-by-sex-age}
knitr::kable(
  descriptives$by_sex_age,
  caption = "Descriptive statistics by sex within age band."
)
```

