---
title: "Video Games, Self-Concept, and the Parent–Child Relationship"
subtitle: "Analysis report"
author: "Taren Sanders"
date: today
# The docx is the reviewed artifact; the html is a quick-preview convenience
# (first so the VS Code Quarto preview uses it). Both render in the pipeline.
format:
  html:
    toc: true
    number-sections: true
    embed-resources: true
  docx:
    toc: true
    number-sections: true
execute:
  echo: false
  warning: false
  message: false
---

```{r setup}
library(targets)
library(ggplot2)
library(knitr)

# Quarto renders from the document's directory, so point tar_read() at the
# project-root store explicitly.
store <- here::here("_targets")
```

# Methods

- Data: LSAC B cohort at ages 10/11, 12/13, and 14/15 (waves 6–8). The K cohort is excluded because its age-14 wave (2012) did not have the video game time item.

## Key Variables

- Video game use: reported minutes of electronic games in an average week, converted to hours/week.
- Self-concept: self-reported SDQ. The total difficulties score is the primary outcome, and internalising (emotional + peer) and externalising (conduct + hyperactivity) composites are for secondary analyses. The Parent 1-reported SDQ total is also kept for an informant-robustness check.
- Parental warmth (the moderator): Parent 1's score on the 6-item Parental Warmth Scale (1–5). Parent 2's score is retained for a both-parents sensitivity model. It's not the primary model because so much of Parent 2's data is missing.
- Covariates: sex, socioeconomic position, and age in months, all at the age 10/11 baseline.
- Cleaning: LSAC missing-data codes recoded to missing. Video game hours more than 4 SD above the within-wave mean are winsorised (that is, set to the 4 SD value). That is different from the proposal (which said set to NA), but I think makes more sense. The flagged values look like real heavy gamers overstating, so we keep them but cap their influence. I'll also test setting them to missing as a sensitivity analysis. SES and warmth are not screened — there are only a few extreme SES values and these seem plausible given the size of LSAC, and low warmth scores are likely to be real given that these are likert scale items.

# Data and derivations

## Sample and missingness per wave

```{r tbl-missingness}
knitr::kable(
  tar_read(missingness_table, store = store),
  caption = "Non-missing observations (% of children in the wave file) for each study variable, after cleaning. N is the number of children with any data at that wave."
)
```

## Distributions

```{r plot-distributions, fig.cap="Distributions of the three core study variables at each age band (after cleaning).", fig.width=8, fig.height=7}
tar_read(distributions_plot, store = store)
```

## Video game hours: outlier tail {#sec-vg-tail}

Values more than 4 SD above the within-wave mean are winsorised to 4 SD from mean. The flagged values are shown below before windsorising.

```{r tbl-vg-tail}
knitr::kable(
  tar_read(vg_tail_table, store = store),
  caption = "Unscreened video game hours per week: within-wave summary and all values above the 4 SD flagging threshold."
)
```

::: {.callout-note title="Checkpoint decisions (Step 1)"}
- Do the derivations (hours/week conversion, SDQ composites, within-wave
  z-scores) look right?
- Decided 2026-07-27: flagged video game hours are winsorised to the 4 SD
  threshold (set-to-missing kept as a sensitivity analysis); SES and
  warmth are not outlier-screened.
- Inclusion criterion flagged for the Step 2 checkpoint: FIML under MAR
  (recommended) vs the proposal's complete-case criterion.
:::
